---
title: "Chess Ratings"
description: |
  Analysis of chess ratings
author: Cole Parra
date: September 30, 2026
format: html
excecute:
  message: false
  warning: false
---

## Introduction

The International Chess Federation, FIDE, releases the ratings for chess players each month. The highest title a chess player can earn is Grandmaster. This Tidy Tuesday analysis uses the ratings from September 2025. My analysis will see which countries have the most Grandmasters.

## Analysis

```{r}
library(tidyverse)

fide_ratings_september <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-09-23/fide_ratings_september.csv')
```

```{r}
fide_ratings_september |>
  filter(title == "GM") |>
  count(fed, sort = TRUE) |>
  slice_head(n = 10) |>
  ggplot(aes(x = n, y = fct_reorder(fed, n))) + 
  geom_col() + 
  geom_text(aes(label = n), hjust = 1.3) +
  labs(
    x = "# of Grandmasters",
    y = "Country",
    title = "Countries with most Grandmasters"
  )
  
```

Germany, the USA, and India are the countries with the most chess Grandmasters, having 77, 75, and 67, respectively. After the top 4 countries, which includes those named and Russia, there is a fairly steep drop off from 63 to 51 Grandmasters, which is Ukraine. Out of the top 10, there are 7 European countries, with the USA, India, and Israel being the only non-European countries. Numbers 5-10 are relatively close together, with each country only having a couple less Grandmasters then the one before it.

## Sources

Data from \[TidyTuesday, September 23, 2025\](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-09-23/readme.md),
curated by Jessica Moore.

Original source: FIDE (International Chess Federation). (2025). \*Standard
Rating List, September 2025\*. https://ratings.fide.com/download_lists.phtml
